---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.13.0
  kernelspec:
    display_name: R 4.1
    language: R
    name: ir41
---

# Setup

```{r tags=c()}
library(here)

source(here::here("Processing","helpers.R") )

seed_num <- 42L


export.path_graphs <- here::here("Figure_Creation","Figures")
dir.create(file.path(export.path_graphs), showWarnings = FALSE,recursive=TRUE)

export.path_objects <- here::here("Processing","results","Objects")
```

```{r}
cell_cycle_cols <- c(G1="#FFC312",G2M="#de5421",S="#12CBC4")
```

```{r}
stromal_cell_type_colors <- c("Septoclasts"="#E64B35FF","OB"="#c2a451","dpMSC1"="#638475","dpMSC2"="#90E39A","mpMSC"="#42c2f5","pMSC"="#00A087FF")
prismatic::color(stromal_cell_type_colors)
```

```{r}
stromal_cell_type_colors_reduced <- c("OB"="#c2a451","dpMSC"="#638475","mpMSC"="#42c2f5","aEC"="#8491B4FF","mpEC"="#4DBBD5FF","bmEC"="#00A087FF","pEC"="darkred")
prismatic::color(stromal_cell_type_colors_reduced)
```

```{r}
merged_cell_type_colors <- c("pvMSC"="#f0a500","Septoclasts"="#E64B35FF","Bone Cells"="#c2a451","aEC"="#8491B4FF","mpEC"="#4DBBD5FF","bmEC"="#00A087FF")
prismatic::color(merged_cell_type_colors)
```

# Load Data


## Full Septoclast Dataset

```{r}
seurat_septoclasts <- readRDS(paste0(export.path_objects,"/Septoclasts_clustered.Rds"))
```

```{r}
DimPlot(seurat_septoclasts,label=TRUE)
```

## Focus on Septoclasts

```{r}
seurat_stromal_subset <- readRDS(paste0(export.path_objects,"/Septoclasts_subset.Rds"))
```

```{r}
unique(seurat_stromal_subset@meta.data$final_cell_type)
DimPlot(seurat_stromal_subset,label=TRUE)
```

## Full metaphyseal Dataset

```{r}
seurat_stromal_all <- readRDS(paste0(export.path_objects,"/Mp_merged.Rds"))

DimPlot(seurat_stromal_all)
```

## Focus on metaphyseal EC and MSC

```{r}
seurat_stromal <- readRDS(paste0(export.path_objects,"/Mp_merged_subset.Rds"))

DimPlot(seurat_stromal,label=TRUE)
```

## Full merged Dataset

```{r}
seurat_merged_all <- readRDS(paste0(export.path_objects,"/Integrated.Rds"))

DimPlot(seurat_merged_all,group.by="final_cell_type")
```

## Merged Focus on metaphyseal EC, Bone Cells, MSC and Septoclasts

```{r}
seurat_all_integrated <- readRDS(paste0(export.path_objects,"/Integrated_focused_subset_withBoneCells.Rds"))

DimPlot(seurat_all_integrated)
```

# Main Figures

<!-- #region tags=[] -->
## Figure 2b
<!-- #endregion -->

```{r}
Figure_2b <- UMAP_Plot(seurat_stromal_subset,
          pt.size=0.5,
          reduction="umap", 
          group.by="final_cell_type",
          split.by=NA,
          facet_labels=NA,
          border_col=NA,
          colors=stromal_cell_type_colors,
          legend=FALSE,
          tick_labels=FALSE,
          n.row=1,
          n.col=NA,
          x.axis_label=TRUE,
          y.axis_label=TRUE,
          x.axis_line=TRUE,
          y.axis_line=TRUE)

Figure_2b

ggsave(here::here(export.path_graphs,"Figure_2b.pdf"),Figure_2b, width=5, height=5, limitsize=F)
```

## Figure 2c

```{r}
Figure_2c <- Feature_Plot(seurat_stromal_subset,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("H2B-GFP","Pdgfra","Pdgfrb","Esm1","Postn","Col22a1","Mmp9","Fabp5","Top2a"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=3,
                          n.col=3,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

Figure_2c

ggsave(here::here(export.path_graphs,"Figure_2c.pdf"),Figure_2c, width=15, height=10, limitsize=F)
```

## Figure 2d + e + f

```{r tags=c()}
library(monocle)
cds_stromal <- readRDS(file=paste0(export.path_objects,"/Monocle2_Septoclast_subset.Rds"))
```

```{r}
combined_monocle_trajectory <- plot_cell_trajectory(cds_stromal, color_by = "clustername",theta = 180,show_branch_points=FALSE) + scale_color_manual(values = stromal_cell_type_colors, name = "cluster") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

combined_monocle_trajectory_split <- plot_cell_trajectory(cds_stromal, color_by = "clustername",theta = 180,show_branch_points=FALSE) + 
scale_color_manual(values = stromal_cell_type_colors, name = "cluster") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) + facet_wrap(~clustername)


combined_monocle_trajectory_genes <- plot_grid(
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Mmp9",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Esm1",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Postn",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Col22a1",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    nrow=2,ncol=2
    )


ggsave(here::here(export.path_graphs,"Figure_2d.pdf"),combined_monocle_trajectory, width=5, height=5, limitsize=F)
ggsave(here::here(export.path_graphs,"Figure_2e.pdf"),combined_monocle_trajectory_split, width=10, height=10, limitsize=F)
ggsave(here::here(export.path_graphs,"Figure_2f.pdf"),combined_monocle_trajectory_genes , width=10, height=10, limitsize=F)


combined_monocle_trajectory_genes
```

## Figure 3a

```{r}
DE_table <- FindMarkers(object = seurat_stromal_subset, ident.1= "Septoclasts",ident.2=NULL,only.pos = FALSE, min.diff.pct = -Inf, min.pct = 0, logfc.threshold = log2(x = 1), test.use = "wilcox",assay="RNA",verbose = TRUE, slot = "counts")
```

```{r}
DE_table_df <- DE_table %>% rownames_to_column(var="gene")
DE_table_df$sig <- DE_table_df$p_val_adj < 0.01 & abs(DE_table_df$avg_log2FC) >=  log2(2)

DE_table_df$sig <- replace_na(DE_table_df$sig,"FALSE")
```

```{r}
######CellScatter#####
average_expression <- (AverageExpression(seurat_stromal_subset, return.seurat = FALSE,assay="RNA",slot="data"))$RNA %>% as.data.frame() %>% rownames_to_column(var="gene")

goi <- c("Mmp9","Fabp5","Mmp13","Mmp14")

subset_matrix <- as.matrix(average_expression[,c("OB", "pMSC","mpMSC","dpMSC1","dpMSC1")])
```

```{r}
average_expression$cluster_mean <- rowMeans(subset_matrix, na.rm=FALSE)
average_expression <- average_expression %>% dplyr::left_join(DE_table_df,by="gene") %>% dplyr::select(gene,Septoclasts,cluster_mean,sig)
average_expression$sig <- replace_na(average_expression$sig,"FALSE")
```

```{r}
scatter_Plot <- ggplot(average_expression,aes(x=log2(1+ Septoclasts),y=log2(1+cluster_mean),label=gene,color=sig ) ) + geom_point(alpha=0.8) + theme_cowplot()  +
  scale_colour_manual(labels = c("p >= 0.01", "p < 0.01"),
                      name = "Significance",values=c("#F00F07","#0B7A79")) +
  guides(colour = guide_legend(override.aes = list(size=2)))+
  geom_segment(aes(x = 0, y = 0, xend = 6, yend = 6),
               colour = "grey")+labs(y= "log2(1+ Stromal Cells mean expression)", x = "log2(1+ Septoclasts mean expression)")+
  scale_x_continuous(expand = c(0.01, 0.01),limits = c(0,6)) +
  scale_y_continuous(expand = c(0.01, 0.01),limits = c(0,6))+
  geom_segment(aes(x = 0,xend=5, y = 1,yend=6),
            colour = "grey70",
            linetype = 2) +
  geom_segment(aes(x = 1,xend=6, y = 0,yend=5),
            colour = "grey70",
            linetype = 2) +
  geom_text(aes(x = 4, y = 5.5, label = ""),
            angle = 56) +
  geom_text(aes(x = 5.5, y = 4, label = ""),
            angle = 56)+
  geom_label_repel(data = average_expression %>% dplyr::filter(gene %in% goi),aes(label = gene),
                   
                   force = 1 )
ggsave(here::here(export.path_graphs,"Figure_3a.pdf"),scatter_Plot , width=5, height=5, limitsize=F)
scatter_Plot
```

## Figure 3c

```{r tags=c()}
library(ComplexHeatmap)
library(circlize)
average_expression <- (AverageExpression(seurat_stromal_subset, return.seurat = FALSE,assay="RNA",slot="data"))$RNA
average_expression_subset <- average_expression[c("Fabp5","Mmp9","Mmp13","Mmp14","Mmp11","Ece1","Adam19"),]
average_expression_subset_scaled <-  (scale(t(average_expression_subset)))
```

```{r}
col_fun = colorRamp2(breaks=c(-2, 0, 2),colors= c("grey65","white","deepskyblue3"))
prismatic::color(col_fun(-2))
```

```{r}

Figure3c <- Heatmap(average_expression_subset_scaled,
        name="Z-Score", 
        show_row_names = TRUE, 
        show_column_names = TRUE,
        row_order=c("Septoclasts","dpMSC1","dpMSC1","mpMSC","OB","pMSC"),
        heatmap_legend_param=list(direction="vertical",at = c(-2, 0, 2),title_position = "leftcenter-rot"),
        col=col_fun,
        column_order= c("Fabp5","Mmp9","Mmp13","Mmp14","Mmp11","Ece1","Adam19"), 
        column_names_rot = 90, 
        column_names_side = "top",
        row_names_side="left",
        border=FALSE)

pdf(file=here::here(export.path_graphs,"Figure_3c.pdf"),width=5,height=5)
draw( Figure3c , heatmap_legend_side = "right")
dev.off()

draw( Figure3c , heatmap_legend_side = "right")
```

## Figure 4b

```{r}
Figure_4b <- UMAP_Plot(seurat_stromal,
          pt.size=0.5,
          reduction="umap", 
          group.by="final_cell_type",
          split.by=NA,
          facet_labels=NA,
          border_col=NA,
          colors=stromal_cell_type_colors_reduced,
          legend=TRUE,
          tick_labels=FALSE,
          n.row=1,
          n.col=NA,
          x.axis_label=TRUE,
          y.axis_label=TRUE,
          x.axis_line=TRUE,
          y.axis_line=TRUE)

Figure_4b

ggsave(here::here(export.path_graphs,"Figure_4b.pdf"),Figure_4b, width=6, height=5, limitsize=F)
```

## Figure 4c

```{r tags=c()}
library(ComplexHeatmap)
library(circlize)

genes_of_interest <- c("Cxcl12","Adipoq","Esm1","Kitl","Lepr","Ramp2","Aplnr","Cd34","Ramp3","Egfl7","Fabp4","Stab2","Mrc1","Selp","Cd36","Bglap2","Col11a1","Car3","Col22a1","Bglap3","Kcnk2","Alpl","Wif1","Vcan","Olfml2b","Gja4","Egfl8","Gja5","Edn1","Sema3g","Mki67","Top2a","Birc5","Cdca8","Ccna2")

average_expression <- (AverageExpression(seurat_stromal, return.seurat = FALSE,assay="RNA",slot="data"))$RNA

average_expression_subset <- average_expression[genes_of_interest,]
average_expression_subset_scaled <-  t(scale(t(average_expression_subset)))
```

```{r}
col_fun = colorRamp2(breaks=c(-2, 0, 2),colors= c("grey65","white","deepskyblue3"))
prismatic::color(col_fun(-2))
```

```{r}
colnames(average_expression_subset_scaled)
```

```{r}

Figure_4c <- Heatmap(average_expression_subset_scaled,
        name="Z-Score", 
        show_row_names = TRUE, 
        show_column_names = TRUE,
        column_order=c("dpMSC","mpEC","bmEC","OB","mpMSC","aEC","pEC"),
        heatmap_legend_param=list(direction="vertical",at = c(-2, 0, 2),title_position = "leftcenter-rot"),
        col=col_fun,
        row_order= genes_of_interest, 
        column_names_rot = 90, 
        column_names_side = "top",
        row_names_side="right",
        border=FALSE)

pdf(file=here::here(export.path_graphs,"Figure_4c.pdf"),width=5,height=5)
draw( Figure_4c , heatmap_legend_side = "right")
dev.off()

draw( Figure_4c , heatmap_legend_side = "right")
```

## Figure 4d

```{r}
Figure_4d <- Feature_Plot(seurat_stromal,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("Pdgfrb","Emcn","Col22a1","Top2a"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=2,
                          n.col=2,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

ggsave(here::here(export.path_graphs,"Figure_4d.pdf"),Figure_4d, width=6, height=5, limitsize=F)

Figure_4d
```

## Figure 4e

```{r}
Figure_4e <- UMAP_Plot(seurat_all_integrated,
          pt.size=0.5,
          reduction="umap", 
          group.by="final_cell_type",
          split.by=NA,
          facet_labels=NA,
          border_col=NA,
          colors=merged_cell_type_colors,
          legend=FALSE,
          tick_labels=FALSE,
          n.row=1,
          n.col=NA,
          x.axis_label=TRUE,
          y.axis_label=TRUE,
          x.axis_line=TRUE,
          y.axis_line=TRUE)

Figure_4e

ggsave(here::here(export.path_graphs,"Figure_4e.pdf"),Figure_4e, width=5, height=5, limitsize=F)
```

## Figure 4f

```{r tags=c()}
library(ComplexHeatmap)
library(circlize)
average_expression <- (AverageExpression(seurat_all_integrated, return.seurat = FALSE,assay="RNA",slot="data"))$RNA
average_expression_subset <- average_expression[c("Mmp9","Mmp13","Mmp14","Mmp11"),]
average_expression_subset_scaled <-  (scale(t(average_expression_subset)))
```

```{r}
col_fun = colorRamp2(breaks=c(-2, 0, 2),colors= c("grey65","white","deepskyblue3"))
prismatic::color(col_fun(-2))
```

```{r}
colnames(average_expression)
```

```{r}

Figure4f <- Heatmap(average_expression_subset_scaled,
        name="Z-Score", 
        show_row_names = TRUE, 
        show_column_names = TRUE,
        row_order=c("Septoclasts","Bone Cells","pvMSC","aEC","mpEC","bmEC"),
        heatmap_legend_param=list(direction="vertical",at = c(-2, 0, 2),title_position = "leftcenter-rot"),
        col=col_fun,
        column_order= c("Mmp9","Mmp13","Mmp14","Mmp11"), 
        column_names_rot = 90, 
        column_names_side = "top",
        row_names_side="left",
        border=FALSE)

pdf(file=here::here(export.path_graphs,"Figure_4f.pdf"),width=5,height=5)
draw( Figure4f , heatmap_legend_side = "right")
dev.off()

draw( Figure4f , heatmap_legend_side = "right")
```

## Figure 4g

```{r}
data <- GetAssayData(seurat_all_integrated,slot="data",assay="RNA")
data <- t( data[c("Hey1","Heyl"),,drop=FALSE] )
data <- as.data.frame(data)
meta_data <- seurat_all_integrated@meta.data %>% rownames_to_column("cell_name") %>% dplyr::select(cell_name,final_cell_type)
```

```{r}
data_interest <- data %>% tibble::rownames_to_column(var="cell_name") %>% pivot_longer(cols=c("Hey1","Heyl"),names_to="feature",values_to="counts") %>% left_join(meta_data)
data_interest %>% head()
```

```{r}
Figure_4g <- ggplot(data_interest, aes(x=final_cell_type, y=counts,fill=final_cell_type))+ 
      geom_violin(bw=0.15)+ 
      theme_cowplot()+ 
      ylab("Expression Level") + 
      geom_jitter(shape=16, position=position_jitter(0.1),alpha=0.1,color="black")+ 
      labs(x="Cell Type",y="Norm. Expression") + scale_fill_manual(values=merged_cell_type_colors) + facet_wrap(~feature,
  nrow = 2,
  ncol = 1) +
  theme(strip.background =element_blank())+guides(fill=FALSE)
               

ggsave(here::here(export.path_graphs,"Figure_4g.pdf"),Figure_4g, width=5, height=5, limitsize=F)


Figure_4g                                        
```

## Figure 4h

```{r}
Figure_4h <- Feature_Plot(seurat_all_integrated,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("Hey1","Dll1","Dll4","Jag1","Jag2","Heyl","Notch1","Notch2","Notch3","Notch4"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=2,
                          n.col=5,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

Figure_4h

ggsave(here::here(export.path_graphs,"Figure_4h.pdf"),Figure_4h, width=25, height=10, limitsize=F)
```

# Supplementary Figures


## Supplementary Figure 4b

```{r}
color_palette <- hue_pal()(length(unique(seurat_septoclasts@meta.data$final_clusters)))

Supp_Figure_4b <- UMAP_Plot(seurat_septoclasts,
          pt.size=0.5,
          reduction="umap", 
          group.by="final_clusters",
          split.by=NA,
          facet_labels=NA,
          border_col=NA,
          colors=color_palette,
          legend=FALSE,
          tick_labels=FALSE,
          n.row=1,
          n.col=NA,
          x.axis_label=TRUE,
          y.axis_label=TRUE,
          x.axis_line=TRUE,
          y.axis_line=TRUE)

Supp_Figure_4b

ggsave(here::here(export.path_graphs,"Supp_Figure_4b.pdf"),Supp_Figure_4b, width=5, height=5, limitsize=F)
```

## Supplementary Figure 4c

```{r}
SuppFigure_4c <- Feature_Plot(seurat_septoclasts,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("H2B-GFP","Pdgfra","Pdgfrb","Col22a1","Acan","S100a4"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=2,
                          n.col=3,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

SuppFigure_4c

ggsave(here::here(export.path_graphs,"SuppFigure_4c.pdf"),SuppFigure_4c, width=15, height=10, limitsize=F)
```

## Supplementary Figure 4d

```{r tags=c()}
library(ComplexHeatmap)
library(circlize)

genes_of_interest <- c("Mmp9","Sstr2","Cd55","Cotl1","Fabp5","Esm1","Kitl","Adipoq","Lepr","Cxcl12","Wif1","Limch1","Postn","Kcnk2","Alpl","Bglap2","Bglap3","Col22a1","Col11a1","Car3","Top2a","Mki67","Birc5","Ccnb2","Cdk1")

average_expression <- (AverageExpression(seurat_stromal_subset, return.seurat = FALSE,assay="RNA",slot="data"))$RNA
average_expression_subset <- average_expression[genes_of_interest,]
average_expression_subset_scaled <-  t(scale(t(average_expression_subset)))
```

```{r}
col_fun = colorRamp2(breaks=c(-2, 0, 2),colors= c("grey65","white","deepskyblue3"))
prismatic::color(col_fun(-2))
```

```{r}
colnames(average_expression_subset_scaled)
```

```{r}

SuppFigure4d <- Heatmap(average_expression_subset_scaled,
        name="Z-Score", 
        show_row_names = TRUE, 
        show_column_names = TRUE,
        column_order=c("Septoclasts","dpMSC1","dpMSC2","mpMSC","OB","pMSC"),
        heatmap_legend_param=list(direction="vertical",at = c(-2, 0, 2),title_position = "leftcenter-rot"),
        col=col_fun,
        row_order= genes_of_interest, 
        column_names_rot = 90, 
        column_names_side = "top",
        row_names_side="right",
        border=FALSE)

pdf(file=here::here(export.path_graphs,"SuppFigure_4d.pdf"),width=5,height=5)
draw( SuppFigure4d , heatmap_legend_side = "right")
dev.off()

draw( SuppFigure4d , heatmap_legend_side = "right")
```

## Supplementary Figure 4e

```{r tags=c()}
library(monocle)
cds_stromal <- readRDS(file=paste0(export.path_objects,"/Monocle2_Septoclast_subset.Rds"))
```

```{r}
combined_monocle_trajectory_genes <- plot_grid(
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Pdgfrb",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Fabp5",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Car3",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    plot_cell_trajectory(cds_stromal,theta = 180,use_color_gradient=TRUE,markers="Top2a",show_branch_points=FALSE) +
      scale_colour_gradient2(low="grey65",high="deepskyblue3",mid="white",midpoint=0,limits=c(-1,1),oob=squish)+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()),
    nrow=2,ncol=2
    )



ggsave(here::here(export.path_graphs,"SuppFigure_4e.pdf"),combined_monocle_trajectory_genes , width=10, height=10, limitsize=F)


combined_monocle_trajectory_genes
```

## Supplementary Figure 6c

```{r tags=c()}
library(ComplexHeatmap)
library(circlize)

genes_of_interest <- c("Fabp5","Ctsb","Mmp9","Mmp11","Mmp13","Mmp14","Ece1","Adam19","Rhod","Rhoj","Tubb2a","Myh9","Vasp","Fn1","Lama4","Itgb5","Itgav","Actn1","Tln1","Flna")

average_expression <- (AverageExpression(seurat_stromal_subset, return.seurat = FALSE,assay="RNA",slot="data"))$RNA
average_expression_subset <- average_expression[genes_of_interest,]
average_expression_subset_scaled <-  t(scale(t(average_expression_subset)))
```

```{r}
col_fun = colorRamp2(breaks=c(-2, 0, 2),colors= c("grey65","white","deepskyblue3"))
prismatic::color(col_fun(-2))
```

```{r}
colnames(average_expression_subset_scaled)
```

```{r}

SuppFigure6c <- Heatmap(average_expression_subset_scaled,
        name="Z-Score", 
        show_row_names = TRUE, 
        show_column_names = TRUE,
        column_order=c("Septoclasts","dpMSC1","dpMSC2","mpMSC","OB","pMSC"),
        heatmap_legend_param=list(direction="vertical",at = c(-2, 0, 2),title_position = "leftcenter-rot"),
        col=col_fun,
        row_order= genes_of_interest, 
        column_names_rot = 90, 
        column_names_side = "top",
        row_names_side="right",
        border=FALSE)

pdf(file=here::here(export.path_graphs,"SuppFigure_6c.pdf"),width=5,height=5)
draw( SuppFigure6c , heatmap_legend_side = "right")
dev.off()

draw( SuppFigure6c , heatmap_legend_side = "right")
```

## Supplementary Figure 8a

```{r}
SuppFigure_8a <- UMAP_Plot(seurat_merged_all,
          pt.size=0.5,
          reduction="umap", 
          group.by="SAMPLE",
          split.by=NA,
          facet_labels=NA,
          border_col=NA,
          colors=c("darkgreen","darkred","darkblue"),
          legend=TRUE,
          tick_labels=FALSE,
          n.row=1,
          n.col=NA,
          x.axis_label=TRUE,
          y.axis_label=TRUE,
          x.axis_line=TRUE,
          y.axis_line=TRUE)

SuppFigure_8a

ggsave(here::here(export.path_graphs,"SuppFigure_8a.pdf"),SuppFigure_8a, width=5, height=5, limitsize=F)
```

## Supplementary Figure 8b

```{r}
color_palette <- hue_pal()(2)

color_palette <- c(stromal_cell_type_colors,"9"=color_palette[[1]],"10"=color_palette[[2]])

SuppFigure_8b <- UMAP_Plot(seurat_merged_all,
          pt.size=0.5,
          reduction="umap", 
          group.by="final_cell_type",
          split.by=NA,
          facet_labels=NA,
          border_col=NA,
          colors=color_palette,
          legend=TRUE,
          tick_labels=FALSE,
          n.row=1,
          n.col=NA,
          x.axis_label=TRUE,
          y.axis_label=TRUE,
          x.axis_line=TRUE,
          y.axis_line=TRUE)

SuppFigure_8b

ggsave(here::here(export.path_graphs,"SuppFigure_8b.pdf"),SuppFigure_8b, width=5, height=5, limitsize=F)
```

## Supplementary Figure 8c


As can be seen by the marker stainings, the Cluster annotated as pEC, might also contain doublets and not just proliferating EC. The majority however, seem to be cycling Endothelial Cells

```{r}
SuppFigure_8c <- Feature_Plot(seurat_merged_all,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("H2B-GFP","Col22a1","Esm1","Pdgfrb","Emcn","Mmp9"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=2,
                          n.col=3,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

SuppFigure_8c

ggsave(here::here(export.path_graphs,"SuppFigure_8c.pdf"),SuppFigure_8c, width=5, height=5, limitsize=F)
```

## Supplementary Figure 8d

```{r tags=c()}
library(ComplexHeatmap)
library(circlize)

genes_of_interest <- c("Bglap","Bglap2","Col22a1","Car3","Gja4","Cav1","Egfl8","Sema3g","Car7","Sstr2","Mmp9","Fabp5","Egfr","Fbln1","Stab2","Rgs4","Cyp4b1","Flt4","Tnc","Wif1","Postn","Alpl","Kcnk2","Aplnr","Prcp","Ramp3","Ednrb","Apln")

average_expression <- (AverageExpression(seurat_all_integrated, return.seurat = FALSE,assay="RNA",slot="data"))$RNA
average_expression_subset <- average_expression[genes_of_interest,]
average_expression_subset_scaled <-  t(scale(t(average_expression_subset)))
```

```{r}
col_fun = colorRamp2(breaks=c(-2, 0, 2),colors= c("grey65","white","deepskyblue3"))
prismatic::color(col_fun(-2))
```

```{r}
colnames(average_expression_subset_scaled)
```

```{r}

SuppFigure8d <- Heatmap(average_expression_subset_scaled,
        name="Z-Score", 
        show_row_names = TRUE, 
        show_column_names = TRUE,
        column_order=c("Bone Cells","aEC","Septoclasts","bmEC","pvMSC","mpEC"),
        heatmap_legend_param=list(direction="vertical",at = c(-2, 0, 2),title_position = "leftcenter-rot"),
        col=col_fun,
        row_order= genes_of_interest, 
        column_names_rot = 90, 
        column_names_side = "top",
        row_names_side="right",
        border=FALSE)

pdf(file=here::here(export.path_graphs,"SuppFigure_8d.pdf"),width=5,height=5)
draw( SuppFigure8d , heatmap_legend_side = "right")
dev.off()

draw( SuppFigure8d , heatmap_legend_side = "right")
```

## Supplementary Figure 8e

```{r}
SuppFigure_8e <- Feature_Plot(seurat_all_integrated,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("Gja4","Piezo2","Stab2","Mmp9","Kcnk2","Col22a1"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=2,
                          n.col=3,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

SuppFigure_8e

ggsave(here::here(export.path_graphs,"SuppFigure_8e.pdf"),SuppFigure_8e, width=5, height=5, limitsize=F)
```

# Further Figures


## Unfiltered Septoclast Dataset Markers

```{r}
Septoclast_Markers <- Feature_Plot(seurat_septoclasts,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("H2B-GFP","Col22a1","Esm1","Pdgfrb","Emcn","Mmp9"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=2,
                          n.col=3,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

Septoclast_Markers
```

## Merged Subset Notch Expression

```{r tags=c()}
library(ComplexHeatmap)
library(circlize)

genes_of_interest <- c("Dll1","Dll4","Jag1","Notch1","Notch3")

average_expression <- (AverageExpression(seurat_all_integrated, return.seurat = FALSE,assay="RNA",slot="data"))$RNA
average_expression_subset <- average_expression[genes_of_interest,]
average_expression_subset_scaled <-  t(scale(t(average_expression_subset)))
```

```{r}
col_fun = colorRamp2(breaks=c(-2, 0, 2),colors= c("grey65","white","deepskyblue3"))
prismatic::color(col_fun(-2))
```

```{r}
colnames(average_expression_subset_scaled)
```

```{r}

Notch_Expression <- Heatmap(average_expression_subset_scaled,
        name="Z-Score", 
        show_row_names = TRUE, 
        show_column_names = TRUE,
        column_order=c("Bone Cells","aEC","Septoclasts","bmEC","pvMSC","mpEC"),
        heatmap_legend_param=list(direction="vertical",at = c(-2, 0, 2),title_position = "leftcenter-rot"),
        col=col_fun,
        row_order= genes_of_interest, 
        column_names_rot = 90, 
        column_names_side = "top",
        row_names_side="right",
        border=FALSE)

Notch_Expression
```

## Merged Subset Markers

```{r}
Merged_Subset_Markers <- Feature_Plot(seurat_all_integrated,
                          pt.size=0.1,
                          scale.label="Norm. Expression",
                          features=c("Emcn","Mmp9","Col22a1","Pdgfrb"),
                          split.by=NA,
                          legend=TRUE,
                          tick_labels=FALSE,
                          use_assay="RNA",
                          use_slot="data",
                          n.row=2,
                          n.col=2,
                          color_lower="grey85",
                          color_upper="deepskyblue3",
                          max.cutoff="q95",
                          min.cutoff=0,
                          reduction="umap")

Merged_Subset_Markers
```

## All Merged Cell Cycle

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
seurat_merged_all <- CellCycleScoring(seurat_merged_all, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)
colnames(seurat_merged_all@meta.data)
```

```{r}
all_merged_cellcycle <- UMAP_Plot(seurat_merged_all,
          pt.size=0.5,
          reduction="umap", 
          group.by="Phase",
          split.by=NA,
          facet_labels=NA,
          border_col=NA,
          colors=cell_cycle_cols,
          legend=TRUE,
          tick_labels=FALSE,
          n.row=1,
          n.col=NA,
          x.axis_label=TRUE,
          y.axis_label=TRUE,
          x.axis_line=TRUE,
          y.axis_line=TRUE)

all_merged_cellcycle
```

```{r}

```
